Alternatives for Neighborhood Function in Kohonen Maps \thanks{This work was supported by a private funding of Velbazhd Software LLC.}

### 2019-02-21 Alternatives-for-Neighborhood-Function-in-Kohonen-Maps/paper.tex
The proposed equations are given.
Small amount of text was added.

\begin{abstract}
In the field of the artificial intelligence artificial neural networks are one of the most researched topics. Multilayer perceptron has a reputation for the most used type of artificial neural network, but other types such as Kohonen maps, generalized nets\cite{tashev01} or combinations with Kalman filter\cite{alexandrov01} are also very interesting. Proposed by Teuvo Kohonen in the 1980s, self-organizing maps have application in meteorology, oceanography, project prioritization and selection, seismic facies analysis for oil and gas exploration, failure mode and effects analysis, creation of artwork and many other areas. Self-organizing maps are very useful for visualization by data dimensions reduction. Unsupervised competitive learning is used in self-organizing maps and the basic idea is the net to classify input data in predefined number of clusters. When the net has fewer nodes it achieve results similar to K-means clustering. One of the components in the self-organizing maps is the neighborhood function. It gives scaling factor for the distance between one neuron and other neurons in each step. The simplest form of a neighborhood function gives 1 for the closest nodes and 0 for all other, but the most used neighborhood function is a Gaussian function. In this research fading cosine and exponential regulated cosine functions are proposed as alternatives for neighborhood function.

@@

Self-organizing maps or Kohonen Neural Networks (KNNs) are networks with unsupervised training. They are very useful in finding nonlinear dependencies when data are presented in very high dimensional spaces. The projection from the high-dimensional space is usually done in a rectangular lower-dimensional space. The main idea behind KNNs is the organization of unlabeled vectors with particular features in predefined number of groups called clusters. Grid of the self-organizing map is a handy tool for convenient visualization which can reveal different features of the network. SOMs are attractive when there consist of two or more separate regions. The real goal is not to find a perfect clustering but get good idea of the cluster structure.

@@

The network consists of a grid (usually two dimensional) with units. The units are connected to adjacent units with a neighborhood relation. The amount of grid units, which usually varies from a few dozen up to several thousand, gives the accuracy and generalization possibilities of the KNN. The network forms an elastic mesh that folds
as a cloud formed by the input data during training phase. Data samples lying near each other in the input space are mapped into nearby grid units.

@@

\begin{equation} \label{equ01}
f(x) = 
	\begin{cases} 
		cos(x) \\
		\frac{cos(x)}{|x|} \\
	\end{cases}
\end{equation}

@@

\begin{thebibliography}{8}
\bibitem{tashev01}
Tashev, T., Hristov, H.: Modeling of synthesis of information processes with generalized nets. Cybernetics and Information Technologies, \textbf{3}(2), 92--104 (2003) 
\bibitem{alexandrov01}
Alexandrov, A.: AD HOC Kalman filter based fusion algorithm for real-time Wireless Sensor Data Integration. In: FQAS-2015 Proceedings, pp. 151--160. Springer, Heidelberg (2015)
\bibitem{vesanto01}
Vesanto, J., Alhoniemi, E.: Clustering of the Self-Organizing Map. IEEE Transactions on Neural Networks, \textbf{11}(3), 586--600 (2000) 
\bibitem{angelova01}
Angelova, V.: Investigations in the Area of Soft Computing Targeted State of the Art Report. Cybernetics and Information Technologies \textbf{9}(1), 18--24 (2009)
\bibitem{atanasova01}
Atanasova T., Barova M.: Exploratory analysis of Time Series for hypothesize feature values. In: UniTech17 Proceedings,  \textbf{16}(2), 399--403 (2017)
\end{thebibliography}
\end{document}

>>>>

\begin{abstract}
In the field of the artificial intelligence artificial neural networks are one of the most researched topics. Multilayer perceptron has a reputation for the most used type of artificial neural network, but other types such as Kohonen maps, generalized nets\cite{tashev01} or combinations with Kalman filter\cite{alexandrov01,voort01} are also very interesting. Proposed by Teuvo Kohonen in the 1980s, self-organizing maps have application in meteorology, oceanography, project prioritization and selection, seismic facies analysis for oil and gas exploration, failure mode and effects analysis, creation of artwork and many other areas. Self-organizing maps are very useful for visualization by data dimensions reduction. Unsupervised competitive learning is used in self-organizing maps and the basic idea is the net to classify input data in predefined number of clusters. When the net has fewer nodes it achieve results similar to K-means clustering. One of the components in the self-organizing maps is the neighborhood function. It gives scaling factor for the distance between one neuron and other neurons in each step. The simplest form of a neighborhood function gives 1 for the closest nodes and 0 for all other, but the most used neighborhood function is a Gaussian function. In this research fading cosine and exponential regulated cosine functions are proposed as alternatives for neighborhood function.

@@

Self-organizing maps or Kohonen Neural Networks (KNNs) are networks with unsupervised training \cite{schreck01}. They are very useful in finding nonlinear dependencies when data are presented in very high dimensional spaces. The projection from the high-dimensional space is usually done in a rectangular lower-dimensional space. The main idea behind KNNs is the organization of unlabeled vectors with particular features in predefined number of groups called clusters \cite{cottrell01}. Grid of the self-organizing map is a handy tool for convenient visualization which can reveal different features of the network. SOMs are attractive when there consist of two or more separate regions. The real goal is not to find a perfect clustering but get good idea of the cluster structure.

@@

The network consists of a grid (usually two dimensional \cite{wehrens01}) with units. The units are connected to adjacent units with a neighborhood relation. The amount of grid units, which usually varies from a few dozen up to several thousand, gives the accuracy and generalization possibilities of the KNN. The network forms an elastic mesh that folds
as a cloud formed by the input data during training phase. Data samples lying near each other in the input space are mapped into nearby grid units.

@@

\begin{equation} \label{equ01}
f(x) = 
	\begin{cases} 
		cos(x) & +\pi/2 \geq x \leq -\pi/2 \\
		\frac{cos(x)}{|x|} & -\pi/2 < x < \pi/2 \\
	\end{cases}
\end{equation}

@@

The proposed neighbor functions give promising results as it is shown in the experimental section. Stressing the neighboring nodes down and up leads to better separation between the formed clusters. As further research it will be interesting the proposed neighborhood functions to be combined with different distance norms. In some problems Hausdorff distance can be used for example.

\begin{thebibliography}{8}

\bibitem{tashev01}
Tashev, T., Hristov, H.: Modeling of synthesis of information processes with generalized nets. Cybernetics and Information Technologies, \textbf{3}(2), 92--104 (2003)

\bibitem{alexandrov01}
Alexandrov, A.: AD HOC Kalman filter based fusion algorithm for real-time Wireless Sensor Data Integration. In: FQAS-2015 Proceedings, pp. 151--160. Springer, Heidelberg (2015)

\bibitem{voort01}
van der Voort, M., Dougherty, M.,Watson, S.: Combining kohonen maps with arima time series models to forecast traffic flow. Transportation Research Part C: Emerging Technologies, \textbf{4}(5), 307--318 (1996)

\bibitem{schreck01}
Schreck, T., Bernard, J., Landesberger, T. von, Kohlhammer, J.: Visual Cluster Analysis of Trajectory Data with Interactive Kohonen Maps. Information Visualization, \textbf{8}(1), 14--29 (2009)

\bibitem{cottrell01}
Cottrell M., Letremy P., Roy E.: Analysing a contingency table with Kohonen maps: A factorial correspondence analysis. In: Mira J., Cabestany J., Prieto A. (eds) New Trends in Neural Computation. Lecture Notes in Computer Science, pp. 305--311. Springer, Heidelberg (1993)

\bibitem{wehrens01}
Wehrens, R., Buydens, L.: Self- and Super-organizing Maps in R: The kohonen Package. Journal of statistical software, \textbf{21}(5), 1--19 (2007)

\bibitem{vesanto01}
Vesanto, J., Alhoniemi, E.: Clustering of the Self-Organizing Map. IEEE Transactions on Neural Networks, \textbf{11}(3), 586--600 (2000)

\bibitem{angelova01}
Angelova, V.: Investigations in the Area of Soft Computing Targeted State of the Art Report. Cybernetics and Information Technologies \textbf{9}(1), 18--24 (2009)

\bibitem{atanasova01}
Atanasova T., Barova M.: Exploratory analysis of Time Series for hypothesize feature values. In: UniTech17 Proceedings,  \textbf{16}(2), 399--403 (2017)

\end{thebibliography}
\end{document}
